Symmetric Circuits for Rank Gates

### 2018-01-30 sections/transparency-and-unique-extensions.tex
30 Jan

@@

\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
  $n$ and let $g_1, g_2 \in G$. Let $D$ be the circuit formed from $C$ by
  removing every gate $g \in G \setminus\{g_1, g_2\}$ such that $\neg W_t(g,
  g_1) \land \neg W_t(g, g_2)$ (i.e. all those gates in the circuit such that
  there is no path from them to either $g_1$ or $g_2$). Here $W_t$ is the
  relation defined in the proof of Lemma~\ref{}. Let $C'$ be the circuit formed
  from $D$ by adding in a single two-input $\AND$ gate $g'$ and connecting the
  outputs of $g_1$ and $g_2$ to the inputs of $g'$. Moreover, we let this $g'$
  be the single output gate of $C'$.

  It follows that $g_1$ and $g_2$ are syntactically equivalent in $C$ if, and
  only if, $g'$ has unique labels in $C'$. Since the construction of $C'$ from
  $C$ can be completed in polynomial time, the mapping of $(C, (g_1, g_2))$ to
  $(C', g')$ is a reduction.
\end{proof}

@@

\begin{lem}
  There is a polynomial-time reduction from the problem of determining if a
  given pair of gates gates in a given $(\mathbb{B}, \rho)$-circuit are
  syntactically-equivalent to the problem of determining if for a given triple
  $(C, g)$, where $C$ is a circuit of order $n$, $g$ is a gate in $C$, that
  there exists $\sigma \in \sym_n$ and automorphisms $\pi, \pi' \in \aut(C)$
  extending $\sigma$ such that $\pi (g) \neq \pi' (g)$.
  \label{lem:syntactically-equivilent-unique-extensions}
\end{lem}
\begin{proof}
  Let $C$ be a circuit of order $n$ and let $g_1$ and $g_2$ be two gates in $C$.
  Note that for any gate $g$ in $\sigma \in sym_n$, if $\pi, \pi' \in \aut(C)$
  extend $\sigma$ and $\pi_e := \pi'\pi{-1}$ then $\pi(g) \neq \pi'(g)$ if, and
  only if, $\pi_e (g) \neq g$. It follows that there exists $\sigma \in sym_n$
  such that there are automorphisms $\pi, \pi' \in \aut(C)$ extending $\sigma$
  such that $\pi (g) \neq \pi'(g)$ if, and only if, there exists $\pi_e$
  extending the trivial permutation such that $\pi (g) \neq g$.

  Let $C'$ be the circuit constructed from $C$ as in the proof of
  Lemma~\ref{lem:syntactically-equivalent-unique-labels}. We now prove that the
  mapping $(C, (g_1, g_2))$ to $(C', g_1)$ is a reduction. Let $\pi_e$ be a
  function from the gates of $C'$ to the gates of $C'$ that swaps $g_1$ and
  $g_2$ and fixes all other gates. It follows that if $g_1$ and $g_2$ are
  syntactically-equivalent in $C$, then they are syntactically-equivalent in
  $C'$, and so $\pi_e$ is a non-trivial automorphism extending the trivial
  permutation, and $g_1$ does not have unique extensions in $C'$. In the other
  direction, suppose $g_1$ does not have unique extensions in $C'$. Then there
  exists an automorphism $\pi_e \in \aut(C')$ extending the trivial permutation
  and such that $\pi_e(g_1) \neq g_1$. But $g_1$ is a child of the single output
  gate $g'$ (which must be fixed by any automorphism), and the only other child
  of $g'$ is $g_2$. It follows $\pi_e$ swaps $g_1$ and $g_2$, and so $g_1$ and
  $g_2$ are syntactically-equivalent in $C'$. The result follows.
\end{proof}

>>>>

  % Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  % $(\mathbb{B}, \rho)$-circuit of order $n$. We define a relation $\sim$ by
  % induction on height (i.e. distance from an input gate). This relation will be
  % the syntactic-equivalence relation on $G$. It is easy to check the conditions
  % for syntactic-equivalence for gates of height 0 as all such gates are input
  % gates and so are either constant gates or relational gates. Suppose $g$ and
  % $h$ are gates of height greater than zero. We only need to check conditions
  % (i), (iv) and (v) of syntactic-equivalence. It is easy to check (i) and (v).
  % If either condition is false then $g$ and $h$ are not syntacticly equivalent.
  % So suppose $g$ and $h$ are two internal gates such that $\Sigma (g) =
  % \Sigma(h)$ and suppose that if they are both output gates then $\Omega^{-1}(g)
  % = \Omega^{-1}(h)$. It remains to check if $L(g)/{\sim}$ and $L(h)/{\sim}$ are
  % isomorphism equivalent. First check if $H_h /{\sim} = H_g /{\sim}$. If not,
  % $g$ and $h$ are not syntactically equivalent. So suppose $H_h /{\sim} = H_g
  % /{\sim}$. If $g$ and $h$ are symmetric gates then $H_h /{\sim} = H_g/{\sim}$
  % if, and only if, $L(g)/{\sim}$ and $L(h)/{\sim}$ are isomorphism equivalent,
  % so let $g \sim h$. If $g$ and $h$ are not symmetric gates then, since $C$ is
  % transparent, $g$ and $h$ have unique labels. We thus have that $L(g)/{\sim}$
  % and $L(h)/{\sim}$ are isomorphism equivalent if, and only if, $L(g)^{-1}L(h)$
  % acts like an isomorphism. This can be checked by first checking that the
  % function induces a sorted permutation on the universe of $g$ (which can be
  % checked efficiently). This is sufficient as $L(g)(L(g)^{-1}L(h)) = L(h)$.

  % It is easy to see that, given a reasonable binary encoding of a circuit, the
  % above procedure can be implemented so as to run in polynomial time, and that
  % the relation $\sim$ is the syntactic-equivalence relation on the gates of $C$.

@@

\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
  $n$ and let $g_1, g_2 \in G$. Let $D$ be the circuit formed from $C$ by
  removing every gate $g \in G \setminus\{g_1, g_2\}$ such that $\neg W_t(g,
  g_1) \land \neg W_t(g, g_2)$, where $W_t$ is the transitive closure of the $W$
  relation (i.e. we move all those gates in the circuit such that there is no
  path from the gate to either $g_1$ or $g_2$). Let $C'$ be the circuit formed from $D$ by adding in a
  single two-input $\AND$ gate $g'$ and connecting the outputs of $g_1$ and
  $g_2$ to the inputs of $g'$. Moreover, we let this $g'$ be the single output
  gate of $C'$.

  It follows that $g_1$ and $g_2$ are syntactically-equivalent in $C$ if, and
  only if, $g'$ has unique labels in $C'$. Since the construction of $C'$ from
  $C$ can be completed in polynomial time, the mapping of $(C, (g_1, g_2))$ to
  $(C', g')$ is a reduction.
\end{proof}

@@

\begin{lem}
  There is a polynomial-time reduction from the problem of determining if a
  given pair of gates gates in a given $(\mathbb{B}, \rho)$-circuit are
  syntactically-equivalent to the problem of determining if for a given pair
  $(C, g)$, where $C$ is a circuit of order $n$, $g$ is a gate in $C$, that
  there exists $\sigma \in \sym_n$ and automorphisms $\pi, \pi' \in \aut(C)$
  extending $\sigma$ such that $\pi (g) \neq \pi' (g)$.
  \label{lem:syntactically-equivilent-unique-extensions}
\end{lem}
\begin{proof}
  Let $C$ be a circuit of order $n$ and let $g_1$ and $g_2$ be two gates in $C$.
  Note that for any gate $g$ in $\sigma \in sym_n$, if $\pi, \pi' \in \aut(C)$
  extend $\sigma$ and $\pi_e := \pi'\pi{-1}$ then $\pi(g) \neq \pi'(g)$ if, and
  only if, $\pi_e (g) \neq g$. It follows that there exists $\sigma \in sym_n$
  such that there are automorphisms $\pi, \pi' \in \aut(C)$ extending $\sigma$
  such that $\pi (g) \neq \pi'(g)$ if, and only if, there exists $\pi_e$
  extending the trivial permutation such that $\pi (g) \neq g$.

  Let $C'$ be the circuit constructed from $C$ as in the proof of
  Lemma~\ref{lem:syntactically-equivalent-unique-labels}. We now prove that the
  mapping $(C, g_1, g_2)$ to $(C', g_1)$ is a reduction. Let $\pi_e$ be
  bijection from the gates of $C'$ to the gates of $C'$ that swaps $g_1$ and
  $g_2$ and fixes all other gates. It follows that if $g_1$ and $g_2$ are
  syntactically-equivalent in $C$, then they are syntactically-equivalent in
  $C'$, and so $\pi_e$ is a non-trivial automorphism extending the trivial
  permutation, and $g_1$ does not have unique extensions in $C'$. In the other
  direction, suppose $g_1$ does not have unique extensions in $C'$. Then there
  exists an automorphism $\pi_e \in \aut(C')$ extending the trivial permutation
  and such that $\pi_e(g_1) \neq g_1$. But $g_1$ is a child of the single output
  gate $g'$ (which must be fixed by any automorphism), and the only other child
  of $g'$ is $g_2$. It follows $\pi_e$ swaps $g_1$ and $g_2$, and so $g_1$ and
  $g_2$ are syntactically-equivalent in $C'$. The result follows.
\end{proof}
